---
title: "Domestic Migration analysis, April 30 2024"
output: 
  html_document:
    css: Wil_style.css
    theme: default
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
always_allow_html: yes
---


```{r setup1, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)


library(dplyr)
library(gglaplot)
library(ggplot2)
library(data.table)
library(lubridate)
library(tidyr)
library(png)
library(stringr)
library(tidyverse)
library(plotly)
library(sf)
library(scales)
library(htmlwidgets)
library(ggplot2)
library(gglaplot)
library(ggrepel)


data_dir <- 'C:/Migration/Migration_R/DATA/Domestic/copied_from_Qdrive_20240213/'
chart_dir <- 'CHARTS/'

# Disable scientific notation
options(scipen=999)


```


```{r setup2, include=FALSE}

## To create SYA migration data frame

# The code used to create the csv files is the same as that used to create the age groups

input_london_dom_mign_in_SYA <-  fread(paste0(data_dir, "summarised_london_destination_SYA.csv")) %>% 
  data.frame()

input_london_dom_mign_out_SYA <- fread(paste0(data_dir, "summarised_london_origin_SYA.csv")) %>%
  data.frame()

input_london_dom_mign_all_SYA = full_join(input_london_dom_mign_in_SYA, input_london_dom_mign_out_SYA,
                                  by=c("year" = "year", "age" = "age", "gss_out" = "gss_in")) %>%
  data.frame() # NB NEED TO DO FULL JOIN NOT LEFT JOIN !!! 

input_london_dom_mign_all_SYA [is.na (input_london_dom_mign_all_SYA)] <- 0

input_london_dom_mign_net_SYA <- input_london_dom_mign_all_SYA %>%
  mutate(sumvaluenetK = sumvalueinK - sumvalueoutK) %>%
  mutate(sumvaluenet = sumvaluein - sumvalueout) %>%
  data.frame()

sum(input_london_dom_mign_net_SYA$sumvalueK_net)

sums <- input_london_dom_mign_net_SYA %>% 
  select(starts_with("sum")) %>% 
  colSums(.) %>%
  data.frame()

sum_London_dom_mign_net_bySYAandYear <- input_london_dom_mign_net_SYA %>%
  group_by(age, year) %>%
  summarize(totalvaluenetK = sum(sumvaluenetK)) %>%
  mutate(label = if_else(year == max(year), 
                         as.character(age), NA_character_)) %>%
  data.frame()

sum_London_dom_mign_net_byYear <- input_london_dom_mign_net_SYA %>%
  group_by(year) %>%
  summarize(totalvaluenetK = sum(sumvaluenetK)) %>%
   data.frame()

write.csv(input_london_dom_mign_net_SYA, paste0(data_dir, "summarised_london_net_SYA.csv"))

```

```{r setup3, include=FALSE}

## FOR CHART ABOUT TRENDS

# Import data for longitudinal chart 
london_longit_net_wide2 <- fread(paste0(data_dir, 
"london_domMig_net_longit_wide2.csv")) %>% 
  data.frame()  


## FOR CHART JUST FOR 2022

# Import data for cross-sectional chart
input_london_dom_mign_net_SYA <- fread(paste0(data_dir, "summarised_london_net_SYA.csv")) %>%
  data.frame()

```


```{r fig_London_domMig_line_1, fig.height = 5.56, fig.width = 9,  echo=FALSE, warning=FALSE, message=FALSE}

##  Plot net migration by age for 2022

London_dom_mign_SYA_net_2022 <- sum_London_dom_mign_net_bySYAandYear %>%
  filter(year == 2022) %>%
  data.frame()

london_dom_mign_net_age_2022 <- 
  ggplot() +
  theme_gla() +
  geom_col(data = London_dom_mign_SYA_net_2022, aes(x = age, y = totalvaluenetK, group = 1), width = 0.5, colour = '#943fa6') +
      geom_hline(yintercept = 0) +
    theme(axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  theme(legend.title=element_text(size=14, family = "Arial"),
        legend.text=element_text(size=11, family = "Arial"),
        axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
   # scale_y_continuous(labels = label_number(suffix = "K", prefix = "")) +
 scale_x_continuous(limits = c(0, 95), breaks = seq(0, 90, 10))  #set breaks on x-axis
  labs(title= 'Net domestic migration by single year of age, London, 2022', 
       caption = paste0(""))
london_dom_mign_net_age_2022

# ggsave (filename = (paste0(chart_dir, "DomMig_Net_flows_age_group_update_noK.png")),
#          plot = Net_flows_age_group_update,
#          device = "png",
#          dpi=600,
#          width = 9,
#          height = 5.56,
#          units = "in")




```

  select(-comboflow) %>%
  data.frame()
# --------------------------------------------------------

# Select those who move between zones in order to calculate net migration between zones

withinLondon_OuterToInner <- withinLondon_primary_agg %>%
  filter(combo == "0 1") %>%
  mutate(comboflow_Inward = comboflow) %>%
  select(-comboflow) %>%
  data.frame()

withinLondon_InnerToOuter <- withinLondon_primary_agg %>%
  filter(combo == "1 0") %>%
  mutate(comboflow_Outward = comboflow) %>%
  select(-comboflow) %>%
  data.frame()

withinLondon_net <-  withinLondon_OuterToInner %>%
  left_join(withinLondon_InnerToOuter, by="age") %>% 
  mutate(net_InnerOuter = comboflow_Outward- comboflow_Inward) %>%
  data.frame()

# ------------------------------------------------------

## 2) Create dataframe for those moving into London from rest of UK 

input_toLondon_fromRegion <- fread(paste0(data_dir, "london_destination_2002to2022.csv")) %>%
 left_join(boroughcodes, by=c("gss_in"="BoroughCode"))%>%
  group_by(gss_in, age, year, Inner, BoroughName) %>%
  summarise(suminflow = sum(value)) %>%
  mutate(entering_inner = factor(Inner)) %>%
  mutate(type = "inflow_fromRestofUK") %>%
  data.frame()

# Filter for 2022 to get total borough values by age
intoLondon_primary_agg <- input_toLondon_fromRegion %>%
  replace(is.na(.), 0) %>%
  filter(year == 2022) %>%
  group_by(age, entering_inner) %>% 
  summarise(comboflow = sum(suminflow)) %>%
  data.frame()

# Select those moving to inner London
FromUK_ToInnerLondon <- intoLondon_primary_agg %>%
  filter(entering_inner == "1") %>%
  mutate(comboflow_ToInner = comboflow) %>%
  select(-comboflow) %>%
  data.frame()

# Select those moving to outer London
FromUK_ToOuterLondon <- intoLondon_primary_agg %>%
  filter(entering_inner == "0") %>%
  mutate(comboflow_ToOuter = comboflow) %>%
  select(-comboflow) %>%
  data.frame()

#-------------------------------------------------------

## 3) Create dataframe for those moving out of London to rest of UK

input_toRegion_fromLondon <- fread(paste0(data_dir, "london_origin_2002to2022.csv")) %>%
  left_join(boroughcodes, by=c("gss_out"="BoroughCode"))%>%
  group_by(gss_out, age, year, Inner, BoroughName) %>%
  summarise(sumoutflow = sum(value)) %>%
  mutate(leaving_inner = factor(Inner)) %>%
  mutate(type = "outflow_toRestofUK") %>%
  data.frame()

# Filter for 2022 to get total borough values by age
outofLondon_primary_agg <- input_toRegion_fromLondon %>%
  replace(is.na(.), 0) %>%
  filter(year == 2022) %>%
  group_by(age, leaving_inner) %>% 
  summarise(comboflow = sum(sumoutflow)) %>%
  data.frame()

# Select those moving from inner London
ToUK_FromInnerLondon <- outofLondon_primary_agg %>%
  filter(leaving_inner == "1") %>%
  mutate(comboflow_fromInner = comboflow) %>%
  select(-comboflow) %>%
  data.frame()

# Select those moving from outer London
ToUK_FromOuterLondon <- outofLondon_primary_agg %>%
  filter(leaving_inner == "0") %>%
  mutate(comboflow_fromOuter = comboflow) %>%
  select(-comboflow) %>%
  data.frame()

# -----------------------------------------------------------------------

## Now calculate the net movements
# Select only those less than 16 years old.

# First for rest of UK v. Inner London
UK_InnerLondon_net <-  ToUK_FromInnerLondon %>%
  left_join(FromUK_ToInnerLondon, by="age") %>% 
  mutate(net_UKInner = comboflow_fromInner - comboflow_ToInner) %>%
  mutate(Flow = "Inner London to rest of UK") %>%
  filter(age < 18) %>%  
  arrange(age) %>%
  mutate(label = if_else(age == max(age),
                         as.character(Flow), NA_character_)) %>%
  data.frame()

# Now for rest of UK v. Outer London 
UK_OuterLondon_net <-  ToUK_FromOuterLondon %>%
  left_join(FromUK_ToOuterLondon, by="age") %>% 
  mutate(net_UKOuter = comboflow_fromOuter - comboflow_ToOuter) %>%
  mutate(Flow = "Outer London to rest of UK") %>%
  filter(age < 18) %>%  
  arrange(age) %>%
  mutate(label = if_else(age == max(age), 
                       as.character(Flow), NA_character_)) %>%
  data.frame()

# Now those who move within zones (again - as this was done above)
withinLondon_net <-  withinLondon_OuterToInner %>%
  left_join(withinLondon_InnerToOuter, by="age") %>% 
  mutate(net_InnerOuter = comboflow_Outward- comboflow_Inward) %>%
  mutate(Flow = "Inner to Outer London") %>%
  filter(age < 18) %>%  
  arrange(age) %>%
  mutate(label = if_else(age == max(age), 
                         as.character(Flow), NA_character_)) %>%
  data.frame()


```



```{r fig_London_domMig_line_2A, fig.height = 5.56, fig.width = 9,  echo=FALSE, warning=FALSE, message=FALSE}

# Plot net migration by age for zero to 17 year olds, for 2022 only

Net_flows_2022 <- 
      ggplot() +
    theme_gla() +
   geom_line(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, colour = '#6da7de',
                                            label = paste("Age  :", age,
                             "<br>Flow from Outer London to rest of UK",
                             "<br>Net flow: ", round(net_UKOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, label = label), 
                      nudge_x = 3,  nudge_y = 0.1, color = '#6da7de')+
  geom_line(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, size = 1, colour = '#d82222',
                                           label = paste("Age  :", age,
                             "<br>Flow from Inner London to rest of UK",
                             "<br>Net flow: ", round(net_UKInner/1000, digits = 0))), size = 2) +
      geom_label_repel(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, label = label), 
                      nudge_x = 3, nudge_y = +0.2, color = '#d82222')+ 
  geom_line(data= withinLondon_net, aes(x = age, y = net_InnerOuter/1000, size = 1, colour = "#eb861e",
                                         label = paste("Age  :", age,
                             "<br>Flow from Inner to Outer London",
                             "<br>Net flow: ", round(net_InnerOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = withinLondon_net, aes(x = age, y = net_InnerOuter/1000, label = label), 
                      nudge_x = 3, nudge_y = -0.1, color = "#eb861e")+ 
  geom_hline(yintercept = 0) +
    theme(axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  theme(legend.title=element_text(size=14, family = "Arial"),
        legend.text=element_text(size=11, family = "Arial"),
        axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  scale_y_continuous (limits = c (0, 3), breaks = c (0, 0.5, 1.0, 1.5, 2.0, 2.5)) +  #set breaks on y-axis
    scale_x_continuous(limits = c (0, 23), breaks = c (0, 4, 8, 12, 17)) +  #set breaks on x-axis
  scale_color_manual(name="Flow direction:", values=c('#6da7de','#d82222','#eb861e'), guide="none") +
    labs(title= "Net domestic migration, year ending June 2022, by age of children, London", 
       caption = paste0(""))
Net_flows_2022

ggsave (filename = (paste0(chart_dir, "DomMig_net_flows_2022_noK.png")),
         plot = Net_flows_2022,
         device = "png",
         dpi=600,
         width = 9,
         height = 5.56,
         units = "in")

# # Net_flows_2020_int <-
# ggplotly(Net_flows_2020, tooltip = "text") %>%  
#   style(hoverlabel = list(bgcolor = "white")) %>%
#   layout(title= list(x = 0.05, 
#                      text = "<b>Net domestic migration in 2020 by age of children<b>", 
#                      font=list(size = 20, family = "Arial", color = "black", fontface = "bold")),
#           legend=list(title=list(text='Flows between:', font = list(size = 15, family = "Arial", color = "black", fontface = "bold"))), 
#           xaxis = list(title = list(text ='Age in years', 
#                                    font = list(size = 15, family = "Arial", color = "black", 
#                                                fontface = "bold"))),
#           yaxis = list(title = list(text ='Annual net migration', 
#                                    font = list(size = 15, family = "Arial", color = "black", 
#                                                fontface = "bold")))
#                  )
# Net_flows_2020_int


```

```{r fig_London_domMig_line_2A_Summ, fig.height = 5.56, fig.width = 9,  echo=FALSE, warning=FALSE, message=FALSE}

# chart for Summary in report

Net_flows_2022_Summ <- 
      ggplot() +
    theme_gla() +
   geom_line(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, colour = '#6da7de',
                                            label = paste("Age  :", age,
                             "<br>Flow from Outer London to rest of UK",
                             "<br>Net flow: ", round(net_UKOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, label = label), 
                      nudge_x = 3,  nudge_y = 0.1, color = '#6da7de')+
  geom_line(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, size = 1, colour = '#d82222',
                                           label = paste("Age  :", age,
                             "<br>Flow from Inner London to rest of UK",
                             "<br>Net flow: ", round(net_UKInner/1000, digits = 0))), size = 2) +
      geom_label_repel(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, label = label), 
                      nudge_x = 3, nudge_y = +0.2, color = '#d82222')+ 
  geom_line(data= withinLondon_net, aes(x = age, y = net_InnerOuter/1000, size = 1, colour = "#eb861e",
                                         label = paste("Age  :", age,
                             "<br>Flow from Inner to Outer London",
                             "<br>Net flow: ", round(net_InnerOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = withinLondon_net, aes(x = age, y = net_InnerOuter/1000, label = label), 
                      nudge_x = 3, nudge_y = -0.1, color = "#eb861e")+ 
  geom_hline(yintercept = 0) +
    theme(axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  theme(legend.title=element_text(size=14, family = "Arial"),
        legend.text=element_text(size=11, family = "Arial"),
        axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  scale_y_continuous (labels = label_number(suffix="K"), limits = c (0, 3), breaks = c (0, 0.5, 1.0, 1.5, 2.0, 2.5)) +  #set breaks on y-axis
  
    scale_x_continuous(limits = c (0, 23), breaks = c (0, 4, 8, 12, 17)) +  #set breaks on x-axis
  scale_color_manual(name="Flow direction:", values=c('#6da7de','#d82222','#eb861e'), guide="none") 
#+
#    labs(title= "Net domestic migration, year ending June 2022, by age of children, London", 
#       caption = paste0(""))
Net_flows_2022_Summ

ggsave (filename = "C:/Families/Charts_for_Daryl/Summ3_DomMig_net_flows_2022_withK.svg",
         plot = Net_flows_2022_Summ,
         device = "svg",
         dpi=600,
         width = 9,
         height = 5.56,
         units = "in")


```

```{r fig_London_domMig_line_2A_B, fig.height = 5.56, fig.width = 9,  echo=FALSE, warning=FALSE, message=FALSE}

# chart for report

Net_flows_2022_rep <- 
      ggplot() +
    theme_gla() +
   geom_line(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, colour = '#6da7de',
                                            label = paste("Age  :", age,
                             "<br>Flow from Outer London to rest of UK",
                             "<br>Net flow: ", round(net_UKOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, label = label), 
                      nudge_x = 3,  nudge_y = 0.1, color = '#6da7de')+
  geom_line(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, size = 1, colour = '#d82222',
                                           label = paste("Age  :", age,
                             "<br>Flow from Inner London to rest of UK",
                             "<br>Net flow: ", round(net_UKInner/1000, digits = 0))), size = 2) +
      geom_label_repel(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, label = label), 
                      nudge_x = 3, nudge_y = +0.2, color = '#d82222')+ 
  geom_line(data= withinLondon_net, aes(x = age, y = net_InnerOuter/1000, size = 1, colour = "#eb861e",
                                         label = paste("Age  :", age,
                             "<br>Flow from Inner to Outer London",
                             "<br>Net flow: ", round(net_InnerOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = withinLondon_net, aes(x = age, y = net_InnerOuter/1000, label = label), 
                      nudge_x = 3, nudge_y = -0.1, color = "#eb861e")+ 
  geom_hline(yintercept = 0) +
    theme(axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  theme(legend.title=element_text(size=14, family = "Arial"),
        legend.text=element_text(size=11, family = "Arial"),
        axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  scale_y_continuous (limits = c (0, 3), breaks = c (0, 0.5, 1.0, 1.5, 2.0, 2.5)) +  #set breaks on y-axis
    scale_x_continuous(limits = c (0, 23), breaks = c (0, 4, 8, 12, 17)) +  #set breaks on x-axis
  scale_color_manual(name="Flow direction:", values=c('#6da7de','#d82222','#eb861e'), guide="none") 
#+
#    labs(title= "Net domestic migration, year ending June 2022, by age of children, London", 
#       caption = paste0(""))
Net_flows_2022_rep

ggsave (filename = "C:/Families/Charts_for_Daryl/22_E_DomMig_net_flows_2022_noK.svg",
         plot = Net_flows_2022_rep,
         device = "svg",
         dpi=600,
         width = 9,
         height = 5.56,
         units = "in")


```

```{r fig_London_domMig_line_2B, fig.height = 5.56, fig.width = 9,  echo=FALSE, warning=FALSE, message=FALSE}

# Plot net migration by age for zero to 17 year olds, for 2022 only

Net_flows_2022 <- 
      ggplot() +
    theme_gla() +
   geom_line(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, colour = '#6da7de',
                                            label = paste("Age  :", age,
                             "<br>Flow from Outer London to rest of UK",
                             "<br>Net flow: ", round(net_UKOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = UK_OuterLondon_net, aes(x = age, y = net_UKOuter/1000, label = label), 
                      nudge_x = 3,  nudge_y = 0.1, color = '#6da7de')+
  geom_line(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, size = 1, colour = '#d82222',
                                           label = paste("Age  :", age,
                             "<br>Flow from Inner London to rest of UK",
                             "<br>Net flow: ", round(net_UKInner/1000, digits = 0))), size = 2) +
      geom_label_repel(data = UK_InnerLondon_net, aes(x = age, y = net_UKInner/1000, label = label), 
                      nudge_x = 3, nudge_y = +0.2, color = '#d82222')+ 
  geom_line(data= withinLondon_net, aes(x = age, y = net_InnerOuter/1000, size = 1, colour = "#eb861e",
                                         label = paste("Age  :", age,
                             "<br>Flow from Inner to Outer London",
                             "<br>Net flow: ", round(net_InnerOuter/1000, digits = 0))), size = 2) +
     geom_label_repel(data = withinLondon_net, aes(x = age, y = net_InnerOuter/1000, label = label), 
                      nudge_x = 3, nudge_y = -0.1, color = "#eb861e")+ 
  geom_hline(yintercept = 0) +
    theme(axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  theme(legend.title=element_text(size=14, family = "Arial"),
        legend.text=element_text(size=11, family = "Arial"),
        axis.text.x=element_text(color = "black", 
                                 size=11, family = "Arial"),  #, angle=30, vjust=.8, hjust=0.8
        axis.text.y=element_text(color = "black", size=11, family = "Arial")) +
  scale_y_continuous (labels = label_number(suffix="K"), limits = c (0, 3), breaks = c (0, 0.5, 1.0, 1.5, 2.0, 2.5)) +  #set breaks on y-axis
    scale_x_continuous(limits = c (0, 23), breaks = c (0, 4, 8, 12, 17)) +  #set breaks on x-axis
  scale_color_manual(name="Flow direction:", values=c('#6da7de','#d82222','#eb861e'), guide="none") +
    labs(title= "Net domestic migration, year ending June 2022, by age of children, London", 
       caption = paste0(""))
Net_flows_2022

ggsave (filename = (paste0(chart_dir, "DomMig_net_flows_2022_withK.png")),
         plot = Net_flows_2022,
         device = "png",
         dpi=600,
         width = 9,
         height = 5.56,
         units = "in")


```
